\huge{Sexual selection improves population fitness: a meta-analysis}

### 2018-11-30 Response_to_Nature_Comms.Rmd

Additionally, the reviewers comment that male attractiveness may not be related to population fitness is correct. Because male attractiveness may be positively or negatively correlated with average fitness (although in many condition-dependent traits a positive association is assumed) we have re-classified *Male Attractiveness* as ambiguously related to fitness, rather than indirect. All subsequent models, statistics and figures now take into account this re-classification. However, given the small sample size for male attractiveness effect sizes (n = 6), **we find only marginal changes in the results and no key changes in the key findings of the study*.**

>*Secondly, looking at Figure 1, one notices apparent anomalies, for example significantly positive slope for male attractiveness, based on studies with average effect size close to zero. I think this (and few others) surprising estimates may result from fitting random terms across all trait types (fitted as a fixed factor, second model); I guess fitting interaction (random slopes) would not be feasible for some categories including few data points, but some of them could easily be pooled in wider categories.*

The reviewer raises questions of the analysis that produce seemingly suprising results. We agree that some of *significant* predictions in Figure 1 (whereby 95 % CIs do not overlap zero) appear surprising based on the distribution on effect sizes. In regards to the trait "male attractiveness", the 95 % CIs do not overlap zero. This is the case even though the higher weighted studies are very close to zero. However, this seeming anomoly can be explained by various reasons.

Firstly, althought the effect size prediction for the trait has 95 % CIs not overlapping zero, the overall magnitude of the effect size is small (see Table exerpt), and the Bayes Factor (an estimate of the likelihood that the effect size is actually positive) is also small. The full table for all predicted effect sizes used in Figure 1 can be seen in **S8**. Below we provide the exerpt of this table and forest plot for "Male attractiveness".

Secondly, the effect size predictions were made using a single model that incorperated all 459 effect sizes, thus model predictions for each trait have narrower confidence intervals as the fitted values do not incorperate measurement error. In regards to our analysis, this means that the predictions that appear in Figure 1 are sourced from using ``fitted.brmsfit`` as opposed to ``predict.brmsfit``. Given the reviewers comments we have clarified how we obtained model predictions used in our results. Additionally, some meta-analyses investigating the effect of a variable on more than one outcome model each outcome (in this case the *fitness component*) seperately. Arguably, this method provides more variable model estimates as models will always have lower power due to reduced sample sizes and the incorperation of measurement error in the predictions. We have added a table to the Supplementary Material where models are independently conducted on all fitness components where n > 3 (**Table S9**).

### 2018-11-30 Manuscripts/Manuscript_MetaAnalysis.Rmd
Furhter changes based on reviewers comments

\newpage
```{r, echo = F, warning=F, message = F}
library(knitr)
library(kableExtra)
library(dplyr)
```
\phantomsection

@@

Conversely, sexual selection can decrease population fitness if male sexually-selected traits are negatively genetically correlated with female fitness, producing intralocus \hl{sexual conflict} [@RN187; @RN146; @RN237; @RN147; @Harano_2010; @Plesnar_2014]. Additionally, sexual selection frequently favours phenotypes that reduce population fitness but benefit the individuals expressing them, such as harassment or infanticide by mate-seeking males [@RN6; @RN267; @RN37], as well as investment in costly sexual signals and weaponry at the expense of parental care: \textit{inter}locus sexual conflict [@RN314; @RN276; @RN315]. Given these conflicting theoretical expectations and empirical results, it remains unclear whether sexual selection tends to have a net benefit or cost to population fitness [@RN42; @RN25; @RN116; @RN95].

@@

We retrieved 459 effect sizes from 65 studies. 94 effect sizes were collected from populations evolving under stressful conditions, while 335 were measured for those evolving in benign conditions. 189 of the effect sizes came from measurements made on males, 219 on females, and the remaining 51 from measurements of a mixed-sex sample of individuals. Most effect sizes in our dataset came from studies that manipulated sexual selection by completely removing it in one treatment via enforced random monogamy ($n$ = 241); other effect sizes ($n$ = 218) derived from alternative manipulations, such as changing the adult sex ratio. In total, we obtained effect sizes for 20 different fitness traits, with reproductive success ($n$ = 156) and offspring viability ($n$ = 56) being the most commonly-measured traits. We classified 216 effect sizes as direct measures of fitness, 147 as indirect, and the remaining 96 effect sizes as ambiguous (see Methods, \autoref{forest_plot} and \href{https://justincally.github.io/SexualSelection/#inclusionexclusion_criteria}{Table S1}). Specifically, direct measures were components of reproduction or long-term viability (*e.g.*, reproductive success, offspring viability, and extinction rate), indirect measures were traits often used as a proxy of fitness but that do not directly measure aspects of success in reproduction or population viability (*e.g.*, lifespan, mating success and ejaculate quality/production); and ambiguous measures were those reported to have an unclear or variable association with fitness (*e.g.*, body size, mating duration, and early fecundity). \href{https://justincally.github.io/SexualSelection/#tables_of_sample_sizes}{Tables S4-S5} give a detailed description of our dataset.\phantomsection

@@

The grand mean for all effect sizes (\hl{direct, indirect and ambiguous}) was positive (REML $\beta$ = 0.23, 95% CIs: 0.01 to 0.44, $p$ = 0.038; Bayesian $\beta$ = 0.23, 95% CIs: -0.06 to 0.51, $BF_{>0}$ = 18), indicating that sexual selection typically had a net positive effect on the majority of populations and fitness components so far studied. Moreover, the effect sizes associated with the manipulation of sexual selection varied between different fitness traits. Sexual selection had a beneficial effect on most fitness traits, including all of the fitness traits that we classified as being ‘direct’ measurements of population fitness: extinction rate (REML $\beta$ = 0.37, 95% CIs: 0.01 to 0.73; Bayesian $\beta$ = 0.37, 95% CIs: -0.02 to 0.79, $BF_{>0}$ = 30.13; $n$ = 4), offspring viability (REML $\beta$ = 0.20, 95% CIs: 0.042 to 0.36; Bayesian $\beta$ = 0.21, 95% CIs: -0.04 to 0.47, $BF_{>0}$ = 24.89; $n$ = 56) and reproductive success (REML $\beta$ = 0.17, 95% CIs: 0.013 to 0.32; Bayesian $\beta$ = 0.18, 95% CIs: -0.06 to 0.44, $BF_{>0}$ = 15.53; $n$ = 156; \autoref{forest_plot}; \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S6}). Sexual selection significantly reduced two fitness components, namely immunity (REML $\beta$ = -0.49, 95% CIs: -0.72 to -0.26; Bayesian $\beta$ = -0.49, 95% CIs: -0.78 to -0.19, $BF_{>0}$ = 0.0054; $n$ = 35) and body condition (REML $\beta$ = -1.3, 95% CIs: -1.9 to -0.67; Bayesian $\beta$ = -1.3, 95% CIs: -1.9 to -0.63, $BF_{>0}$ = 0.00025, $n$ = 1). \autoref{forest_plot} presents Bayesian estimates for each fitness component and \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S6} contains both Bayesian and REML estimates, with sample sizes and test statistics.  \phantomsection

@@

We found that the sex of the individuals measured (male, female, or a mixture), and the conditions under which the population evolved (stressful or benign) interacted to affect fitness (\hyperlink{page.13}{Table 1, 2}). Sexual selection significantly improved female fitness, and the beneficial effect of sexual selection was significantly stronger for females from populations evolving under stressful conditions (*e.g.* a food source to which they were not well-adapted) than under benign conditions (\autoref{env_sex}a, \hyperlink{page.13}{Table 2}). Sexual selection had a positive but non-significant effect on male fitness, and in contrast to females, fitness benefits were significantly weaker in stressful than benign environments (\autoref{env_sex}a, \hyperlink{page.13}{Table 2}). Consistent with the different consequences of sexual selection for female and male fitness, the mean effect size in mixed-sex samples was non-significantly positive, and there was no significant difference between benign versus stressful conditions (\autoref{env_sex}a, \hyperlink{page.13}{Table 2}). Overall, our results indicate that the positive effect of sexual selection on fitness is greater for females than males, and the difference between the sexes is magnified in stressful environments.

Other moderator variables that we examined had minimal impacts on effect size (\href{https://justincally.github.io/SexualSelection/#meta-analyses_including_many_moderator_variables}{Figure S2, Table S7}). Specifically, effect size did not depend on whether or not the study was conducted blind (\href{https://justincally.github.io/SexualSelection/#other_moderators}{Figure S9}), nor on the number of generations for which the experimental evolution study was run (\href{https://justincally.github.io/SexualSelection/#other_moderators}{Figure S10, S11}).

@@

![Forest plot showing the effect of manipulation of sexual selection, for 20 fitness components from 65 studies. Effect sizes ($n$ = 459) varied between fitness components, and were more often positive than negative. Predicted average values are presented alongside each fitness component; those in bold have 95% confidence intervals that do not cross zero. The estimates presented here are Bayesian predictions, which were slightly more conservative than the REML predictions (Table S6).\label{forest_plot}](Figures/ForestPlot_edit3.pdf)

<!-- Appears to be a problem having two captions -->
\newpage
\nolinenumbers

table1 <- read.csv('table1.csv') %>% 
  mutate(
    Estimate = cell_spec(Estimate, "latex", bold = ifelse(LCI*UCI > 0, T, F)),
    Parameters = cell_spec(Parameters, "latex", bold = ifelse(LCI*UCI > 0, T, F))
  )

@@

>>>>

\newpage
```{r, echo = F, warning=F, message = F}
library(knitr)
library(kableExtra)
library(dplyr)
library(brms)

source("../Tidy_functions_for_brms.R") #Tidy functions for making model tables for brms
```
\phantomsection

@@

Conversely, sexual selection can decrease population fitness if male sexually-selected traits are negatively genetically correlated with female fitness, producing intralocus **sexual conflict** [@RN187; @RN146; @RN237; @RN147; @Harano_2010; @Plesnar_2014]. Additionally, sexual selection frequently favours phenotypes that reduce population fitness but benefit the individuals expressing them, such as harassment or infanticide by mate-seeking males [@RN6; @RN267; @RN37], as well as investment in costly sexual signals and weaponry at the expense of parental care: \textit{inter}locus sexual conflict [@RN314; @RN276; @RN315]. Given these conflicting theoretical expectations and empirical results, it remains unclear whether sexual selection tends to have a net benefit or cost to population fitness [@RN42; @RN25; @RN116; @RN95].

@@

We retrieved 459 effect sizes from 65 studies. 92 effect sizes were collected from populations evolving under stressful conditions, while 337 were measured for those evolving in benign conditions. 189 of the effect sizes came from measurements made on males, 219 on females, and the remaining 51 from measurements of a mixed-sex sample of individuals. Most effect sizes in our dataset came from studies that manipulated sexual selection by completely removing it in one treatment via enforced random monogamy ($n$ = 241); other effect sizes ($n$ = 218) derived from alternative manipulations, such as changing the adult sex ratio. In total, we obtained effect sizes for 20 different fitness traits, with reproductive success ($n$ = 156) and offspring viability ($n$ = 56) being the most commonly-measured traits. We classified 216 effect sizes as direct measures of fitness, 141 as indirect, and the remaining 102 effect sizes as ambiguous (see Methods, \autoref{forest_plot} and \href{https://justincally.github.io/SexualSelection/#inclusionexclusion_criteria}{Table S1}). Specifically, direct measures were components of reproduction or long-term viability (*e.g.*, reproductive success, offspring viability, and extinction rate), indirect measures were traits often used as a proxy of fitness but that do not directly measure aspects of success in reproduction or population viability (*e.g.*, lifespan, mating success and ejaculate quality/production); and ambiguous measures were those reported to have an unclear or variable association with fitness (*e.g.*, body size, mating duration, and early fecundity). \href{https://justincally.github.io/SexualSelection/#tables_of_sample_sizes}{Tables S4-S5} give a detailed description of our dataset.\phantomsection

@@

The grand mean for all effect sizes (**direct, indirect and ambiguous**) was positive (REML $\beta$ = 0.25, 95% CIs: 0.06 to 0.43, $p$ = 0.008; Bayesian $\beta$ = 0.25, 95% CIs: -0.02 to 0.52, $BF_{>0}$ = 32), indicating that sexual selection typically had a net positive effect on the majority of populations and fitness components so far studied. Moreover, the effect sizes associated with the manipulation of sexual selection varied between different fitness traits. Sexual selection had a beneficial effect on most fitness traits, but varied across the three relationships to fitness \autoref{forest_plot}; \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S6}). **Sexual selection elevated fitness for traits that shared an ambiguous relationship to fitness (REML $\beta$ = 0.23, 95% CIs: 0.12 to 0.35; Bayesian $\beta$ = 0.23, 95% CIs: 0.03 to 0.41, $BF_{>0}$ = 55, $n$ = 102) and an indirect relationship to fitness (REML $\beta$ = 0.25, 95% CIs: 0.13 to 0.36; Bayesian $\beta$ = 0.24, 95% CIs: 0.04 to 0.42, $BF_{>0}$ = 68, $n$ = 141). Additionally, sexual selection elevated fitness components directly related to fitness, albeit at a lower magnitude (REML $\beta$ = 0.13, 95% CIs: 0.02 to 0.24; Bayesian $\beta$ = 0.13, 95% CIs: -0.07 to 0.3, $BF_{>0}$ = 14, $n$ = 216). An extensive forest plot with predicted effect sizes for each fitness component is presented in Figure S1 and further detailed in model prediction and individual meta-analyses in** \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S8 and S9}. Sexual selection significantly reduced two fitness components, namely immunity (REML $\beta$ = -0.42, 95% CIs: -0.64 to -0.20; Bayesian $\beta$ = -0.43, 95% CIs: -0.70 to -0.16, $BF_{>0}$ = 0.0021; $n$ = 35) and body condition (REML $\beta$ = -1.2, 95% CIs: -1.8 to -0.63; Bayesian $\beta$ = -1.2, 95% CIs: -1.8 to -0.61, $BF_{>0}$ = 0.00013, $n$ = 1).  \phantomsection

@@

We found that the sex of the individuals measured (male, female, or a mixture), and the conditions under which the population evolved (stressful or benign) interacted to affect fitness (\hyperlink{page.18}{Table 1, 2}). Sexual selection significantly improved female fitness, and the beneficial effect of sexual selection was significantly stronger for females from populations evolving under stressful conditions (*e.g.* a food source to which they were not well-adapted) than under benign conditions (\autoref{env_sex}a, \hyperlink{page.18}{Table 2}). Sexual selection had a positive but non-significant effect on male fitness, and in contrast to females, fitness benefits were significantly weaker in stressful than benign environments (\autoref{env_sex}a, \hyperlink{page.18}{Table 2}). Consistent with the different consequences of sexual selection for female and male fitness, the mean effect size in mixed-sex samples was non-significantly positive, **with a small but significant difference** between benign versus stressful conditions (\autoref{env_sex}a, \hyperlink{page.13}{Table 2}). Overall, our results indicate that the positive effect of sexual selection on fitness is greater for females than males, and the difference between the sexes is magnified in stressful environments. **When the lnRR (log Response Ratio) effect size was used to conduct the meta-analysis, the results corroborate those of Hedges' g in that sexual selection elevates population fitness, with its effect magnified for females evolving in stressful environments (Figure S4 and Table S16).**

Other moderator variables that we examined had minimal impacts on effect size (\href{https://justincally.github.io/SexualSelection/#meta-analyses_including_many_moderator_variables}{Figure S2, Table S10}). Specifically, effect size did not depend on whether or not the study was conducted blind (\href{https://justincally.github.io/SexualSelection/#other_moderators}{Figure S9}), nor on the number of generations for which the experimental evolution study was run (\href{https://justincally.github.io/SexualSelection/#other_moderators}{Figure S10, S11}).

@@

<!-- Appears to be a problem having two captions -->

\newpage

@@

\newpage
\phantomsection
\nolinenumbers

#Tables and Figures

```{r, echo = F}
model.complete2 <- readRDS('../data/model.complete2.rds')
brms.complete2 <- readRDS('../data/brms.complete2.rds')

table1 <- data.frame(round(as.numeric(model.complete2[["beta"]]), digits = 2))  %>% rename(Estimate = 1)
table1$SE <- as.numeric(model.complete2[["se"]])
table1$LCI <- as.numeric(model.complete2[["ci.lb"]])
table1$UCI <- as.numeric(model.complete2[["ci.ub"]])
table1$z <- as.numeric(model.complete2[["zval"]])
table1$p <- as.numeric(model.complete2[["pval"]])

brms.model.table <- make_text_summary(brms.complete2) %>% add_significance_stars()
brms.model.table$z <- NA
brms.model.table$p <- NA
brms.model.table <- brms.model.table[1:6,]

table1 <- rbind(table1, brms.model.table %>% select(Estimate, SE = Est.Error, LCI = Q2.5, UCI = Q97.5, z, p)) %>% tibble::rownames_to_column("Parameters")
table1$Parameters <- c("Intercept", "Both Sexes", "Female Sex", "Stressed Environment", "Both Sexes x Stressed Environment", "Female Sex x Stressed Environment", "Intercept", "Both Sexes", "Female Sex", "Stressed Environment", "Both Sexes x Stressed Environment", "Female Sex x Stressed Environment")

options(knitr.kable.NA = ' ')

table1 <- table1 %>% 
  mutate(
    Estimate = cell_spec(Estimate, "latex", bold = ifelse(LCI*UCI > 0, T, F)),
    Parameters = cell_spec(Parameters, "latex", bold = ifelse(LCI*UCI > 0, T, F))
  )

\newpage

![The effect sizes used in this meta-analysis ($n$ = 459) were grouped into either direct, indirect or ambiguous measures of fitness. Overall, effect sizes were more often positive than negative. Predicted average values are presented as a diamond for each fitness-relationship category. The estimates presented here are from REML models with the grand mean across all effect sizes ($\beta$ = 0.25) shown as the blue dotted line. Predictions from both Bayesian and REML models can be found in Table S7).\label{forest_plot}](Figures/forest.plot.class.pdf)
